---
title: "Stock Price Analysis Pipeline"
output: html_document
---

Setup in RStudio (tutorial: <https://www.youtube.com/watch?v=TJmNvfhLCoI>):

-- Look at the above menu bar. Choose the little gear on top for **Preview in View Panel**, then click **Knit to HTML** or **Preview**

-- Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

-- The code is executable in the editor panel below with the green arrow.

Libraries

```{r setup, include=FALSE}
library(quantmod)
library(dplyr)
library(fitdistrplus)
```

------------------------------------------------------------------------

# Functions

## Peak detection

```{r}
find_peaks <- function(data) {
  price <- Cl(data)
  prev  <- lag(price, 1)
  nexts <- lead(price, 1)
  peaks <- (price > prev) & (price > nexts)
  peaks[is.na(peaks)] <- FALSE
  data[peaks]
}
```

## Trough detection

```{r}
find_troughs <- function(data) {
  price <- Cl(data)
  prev  <- lag(price, 1)
  nexts <- lead(price, 1)
  trough <- (price < prev) & (price < nexts)
  trough[is.na(trough)] <- FALSE
  data[trough]
}
```

## Biggest drawdown

```{r}
biggest_drop <- function(clean_price) {
  price <- Cl(clean_price)
  run_peak <- cummax(price)
  dd <- price - run_peak

  end_i <- which.min(dd)
  start_i <- which.max(price[1:end_i])

  start <- as.numeric(price[start_i])
  end   <- as.numeric(price[end_i])

  data.frame(
    start_date = index(price)[start_i],
    end_date   = index(price)[end_i],
    fluct_pct   = (start - end) / start * 100
  )
}
```

## Identify Bull segment

```{r}
bull_segment <- function(clean_price) {
    price <- Cl(clean_price)
  
  if (is.null(price) || NROW(price) < 2) {
    return(NULL)
  }
  
  min_i <- which.min(price)
  
  # guard: no data after min
  if (is.na(min_i) || min_i >= NROW(price)) {
    return(list(
      min_date = index(price)[min_i],
      min_price = as.numeric(price[min_i]),
      peak_after_min_date = NA,
      peak_after_min_price = NA
    ))
  }
  
  after_min <- price[(min_i + 1):NROW(price)]
  
  # extra safety: empty slice
  if (NROW(after_min) == 0) {
    return(list(
      min_date = index(price)[min_i],
      min_price = as.numeric(price[min_i]),
      peak_after_min_date = NA,
      peak_after_min_price = NA
    ))
  }
  
  peak_i_rel <- which.max(after_min)
  
  # guard: which.max can still fail on weird input
  if (length(peak_i_rel) == 0 || is.na(peak_i_rel)) {
    return(list(
      min_date = index(price)[min_i],
      min_price = as.numeric(price[min_i]),
      peak_after_min_date = NA,
      peak_after_min_price = NA
    ))
  }
  
  peak_i <- min_i + peak_i_rel
  
  list(
    min_date = index(price)[min_i],
    min_price = as.numeric(price[min_i]),
    peak_after_min_date = index(price)[peak_i],
    peak_after_min_price = as.numeric(price[peak_i])
  )
}
```

## Fluctuation extractor

```{r}
get_fluctuations <- function(price, min_date, peak_date) {
  segment <- price[paste(min_date, peak_date, sep = "/")]

  prev <- lag(segment, 1)

  fluct_pct <- (as.numeric(segment) - as.numeric(prev)) / as.numeric(prev) * 100
  fluct_pct <- fluct_pct[!is.na(fluct_pct)]

  fluct_pct
}
```

# Demo

```{r}
symbol <- "FIG"
```

## Data loading

```{r}
thisData <- getSymbols(symbol, from = "2025-04-11", auto.assign = FALSE)
```

## Peak / Trough extraction

```{r}
peaks <- find_peaks(thisData)
troughs <- find_troughs(thisData)

# and a combined of these
clean_price <- rbind(peaks, troughs)
clean_price <- clean_price[order(index(clean_price))]
```

## Visualization

```{r}
plot(Cl(thisData), main = paste(symbol," Closing Price"), type = "b")
plot(Cl(clean_price), main = paste(symbol," Clean Closing Prices"), type = "b")
plot(Cl(peaks), main = "Peaks", type = "b")
plot(Cl(troughs), main = "Troughs", type = "b")
```

## Catch the drop

```{r}
recentBottom = biggest_drop(clean_price)
print(recentBottom)
```

## Fluctuation Study for the Bull Segment

```{r}
BullPeriod = bull_segment(clean_price) ## Get the bull segment first
print(BullPeriod)

if (any(is.na(BullPeriod$min_date), is.na(BullPeriod$peak_after_min_date))) {
    warning("Skipping fluctuation calculation: NA dates detected")
    fluc_pct <- NULL
} else {
    fluc_pct <- get_fluctuations(...)
    fluc_pct <- as.numeric(fluc_pct)
  
    # remove NA just in case
    
    fluc_pct <- fluc_pct[!is.na(fluc_pct)]
    
    hist(as.numeric(fluc_pct),
         main = "Bull Segment Fluctuations",
         xlab = "Peak-Trough Fluctuation (%)",
         col = "lightblue",
         breaks = 30)
    
    cat("Worst drop (%):", round(min(fluc_pct), 2), "\n")
}
# ensure numeric



```

## Price Time Series (Remove Continuous Rises/Drops)

```{r}
price <- Cl(clean_price)

fluct_pct <- diff(price) / lag(price) * 100
fluct_pct <- fluct_pct[!is.na(fluct_pct)]
par(mfrow = c(2,1))

plot(price, main = "Clean Price")

hist(fluct_pct,
     breaks = 50,
     main = "Daily % Change Distribution",
     xlab = "% Change",
     col = "lightblue")
```

### Gaussian Fit the % Change

```{r}
fit <- fitdist(fluct_pct, "norm")

mu <- fit$estimate["mean"]
sigma <- fit$estimate["sd"]

hist(fluct_pct,
     breaks = 50,
     probability = TRUE,
     main = "Gaussian Fit",
     xlab = "% Change",
     col = "lightblue")

curve(dnorm(x, mean = mu, sd = sigma),
      add = TRUE,
      col = "red",
      lwd = 2)

cat("Mean (mu):", round(mu, 4), "\n")
cat("Std  (sigma):", round(sigma, 4), "\n")
```

### Get Quantile

```{r}
qnorm(0.05, mean = mu, sd = sigma)
```

### Sample daily return check

```{r}
ret <- dailyReturn(Cl(thisData))

mu <- mean(ret, na.rm = TRUE)
sigma <- sd(ret, na.rm = TRUE)

hist(ret,
     main = paste("Daily Returns of", symbol),
     xlab = "Return",
     col = "lightblue",
     breaks = 50)

legend("topright",
       legend = sprintf("mean = %.4f\nsd = %.4f", mu, sigma),
       bty = "n")
```
